---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r load_packages, include=FALSE}
library(tidyverse)
library(lattice)
```


## Loading and preprocessing the data
```{r reading_data, echo=TRUE}
Data <- read.csv("activity.csv", header=TRUE, sep=",", na.strings = "NA")
```

## What is mean total number of steps taken per day?

```{r median_mean_steps, echo=TRUE}
hist(Data$steps)
mean_steps <- mean(Data$steps, na.rm=TRUE)
median_steps <- median(Data$steps, na.rm=TRUE)
```
The mean of the total number of steps per day is `r mean_steps` and the median `r median_steps`.

## What is the average daily activity pattern?

```{r plot_steps_interval, echo=TRUE}
Dataplot <- Data %>% group_by(interval) %>% summarise(Avg=mean(steps, na.rm=TRUE))
with(Dataplot, plot(interval, Avg, type="l"))
interval <- with(Dataplot, Dataplot[Avg==max(Avg, na.rm=TRUE),]$interval)
Avg <- with(Dataplot, Dataplot[Avg==max(Avg, na.rm=TRUE),]$Avg)
```

The 5-minute interval that contains the maximum number of steps on average across all the days in the dataset is `r interval`, with `r Avg` number of steps.

## Imputing missing values

```{r NA_missing_values, echo=TRUE}
NA_lines <- sum(is.na(Data$steps))

#Will replace NA values by the mean of the steps by inteval
replace.na.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
Data_nna <- Data %>%
                group_by(interval) %>%
                mutate(steps.no.NA=replace.na.mean(steps))
                               
#plot the histogram as before
with(Data_nna, hist(steps.no.NA))

#producing the mean and the median to compare to previous data
mean_steps1 <- mean(Data$steps, na.rm=TRUE)
median_steps1 <- median(Data$steps, na.rm=TRUE)

```

The total number of NAs in the data set are `r NA_lines`
NAs will be replace by the mean of the steps in the respective interval

The mean of the total number of steps per day is `r mean_steps1` and the median `r median_steps1`.  Not significatly different from the initial results.


## Are there differences in activity patterns between weekdays and weekends?

```{r weekday_comparision, echo=TRUE}
#Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

Data_nna$DayType <- factor((weekdays(as.Date(Data_nna$date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")), levels=c(FALSE, TRUE), labels=c('Weekend', 'Weekday'))
         
Dataplot2 <- Data_nna %>% group_by(interval,DayType) %>% summarise(Avg=mean(steps, na.rm=TRUE))

with(Dataplot2, xyplot(Avg~interval | DayType , type="l", layout=(c(1,2))))

```


